Exploring how changes in age at death and cause of death have narrowed the black-white life expectancy gap over forty-five years in the United States.

### 2017-01-05 Code/append_smoothed_results.Rmd
added heatmaps

for(i in 1:dim(paired.ids)[1]) {
    if(list.cod.decomp.tables.smoothed[[i]][1] != "Couldn't compare life tables" & 
       list.cod.decomp.tables.smoothed[[i]][1] != "Encountered age group with no deaths") { 
       # list.cod.marginal.tables.smoothed[[i]] <-  list.cod.decomp.tables.smoothed[[i]] %>%
       #                                            dplyr::group_by(Cause.of.death) %>%
       #                                            dplyr::summarise(C_x_COD = sum(C_xi)) %>%
       #                                            dplyr::mutate(C_x_COD_proportion = C_x_COD/sum(C_x_COD),
       #                                                   adds_to_gap = ifelse(C_x_COD > 0, T, F))

@@

```{r save_specified_files}
save(dat.clean, dat2, BlackWhite, BlackWhiteSmooth, paired.ids, list.age.decomp.tables, list.cod.decomp.tables,
     list.age.decomp.tables5, list.cod.decomp.tables5, list.age.decomp.tables9, list.cod.decomp.tables9,
     list.age.decomp.tables.smoothed, list.cod.decomp.tables.smoothed, list.cod.marginal.tables.smoothed,
     non.zero.stratums, non.zero.stratums.smoothed,
     file = "/Users/corinneriddell/Dropbox/BlackWhiteGap/Data/main_datasets.Rdata")
```

>>>>

for(i in 1:dim(paired.ids)[1]) {
    if(list.cod.decomp.tables.smoothed[[i]][1] != "Couldn't compare life tables" & 
       list.cod.decomp.tables.smoothed[[i]][1] != "Encountered age group with no deaths") { 
       list.cod.marginal.tables.smoothed[[i]] <-  list.cod.decomp.tables.smoothed[[i]] %>%
                                                  dplyr::group_by(Cause.of.death) %>%
                                                  dplyr::summarise(C_x_COD = sum(C_xi)) %>%
                                                  dplyr::mutate(C_x_COD_proportion = C_x_COD/sum(C_x_COD),
                                                         adds_to_gap = ifelse(C_x_COD > 0, T, F))

@@

```{r save_specified_files}
save(dat.clean, dat2, BlackWhite, BlackWhiteSmooth, paired.ids, list.age.decomp.tables, list.cod.decomp.tables,
     list.age.decomp.tables5, list.cod.decomp.tables5, list.age.decomp.tables9, list.cod.decomp.tables9,
     list.age.decomp.tables.smoothed, list.cod.decomp.tables.smoothed, list.cod.marginal.tables.smoothed,
     non.zero.stratums, non.zero.stratums.smoothed, cod.contributions,
     file = "/Users/corinneriddell/Dropbox/BlackWhiteGap/Data/main_datasets.Rdata")
```

### 2017-01-05 Manuscript/Manuscript.Rmd
added heatmaps

```{r figure6, fig.align="center", fig.height=8, fig.width=8, echo=F}
ggplotly(ggplot(subset(BlackWhite, Sex2 == "Female"), aes(y=WBgap.s, x = Year3)) + geom_line(aes(col = State2))
         + facet_wrap(~Census_Region))
```

**Figures 7 and 8 to be added: Trends in the contribution of CVD to the black-white life expectancy gap (in years) by US state**
-make one for males and one for females
-separately for each COD group
-included CIs

>>>>

```{r figure6, fig.align="center", fig.height=8, fig.width=9, echo=F}
ggplotly(ggplot(subset(BlackWhite, Sex2 == "Female"), aes(y=WBgap.s, x = Year3)) + geom_line(aes(col = State2))
         + facet_wrap(~Census_Region) +
           ylab("Black-white life expectancy gap (years)") +
           xlab("Year") +
           geom_text(data = subset(BlackWhite, Year3 == 2013 & Sex2 == "Female"), aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

**Figure 7a: Contribution of cancer to the black-white life expectancy gap for males between 1969 and 2013 by US state, in years**
```{r figure7a, fig.align="center", fig.height=8, fig.width=9, echo=F}
ggplotly(ggplot(subset(cod.contributions, Sex2 == "Male" & Cause.of.death == "Cancers"), aes(x = Year3, y = C_x_COD)) + 
         geom_line(aes(col = State2)) + facet_wrap(~ Census_Region) +
         ylab("Contribution to LE Gap (years)") +
         xlab("Year") +
         geom_text(data = subset(cod.contributions, Year3 == 2013 & Sex2 == "Male" & Cause.of.death == "Cancers"), 
         aes(label = stabbrs), check_overlap = T, size = 2.5)
)

    #theme_grey(base_size = 9) + labs(x = "", y = "") + 
    #scale_x_discrete(expand = c(0, 0)) +
    #scale_y_discrete(expand = c(0, 0)) +
    #theme(legend.position = "none", axis.ticks = element_blank(), axis.text.x = element_text(size = 9*0.8, angle = 330, hjust = 0, colour = "grey50"))

#something is wrong with the percent contribution calculated for RI around 2008 ... hmmph
# ggplotly(ggplot(subset(cod.contributions, Sex2 == "Male" & Cause.of.death == "Cancers"), aes(x = Year3, y = C_x_COD_proportion)) + 
#          geom_line(aes(col = State2)) + facet_wrap(~ Census_Region) +
#          ylab("Contribution to LE Gap (years)") +
#          xlab("Year") +
#          geom_text(data = subset(cod.contributions, Year3 == 2013 & Sex2 == "Male" & Cause.of.death == "Cancers"), 
#          aes(label = stabbrs), check_overlap = T, size = 2.5))
```

**Figure 7b: Contribution of cardiovascular disease to the black-white life expectancy gap for males between 1969 and 2013 by US state, in years**
```{r figure7b, fig.align="center", fig.height=8, fig.width=9, echo=F}
ggplotly(ggplot(subset(cod.contributions, Sex2 == "Male" & Cause.of.death == "Cardiovascular"), aes(x = Year3, y = C_x_COD)) + 
         geom_line(aes(col = State2)) + facet_wrap(~ Census_Region) +
         ylab("Contribution to LE Gap (years)") +
         xlab("Year") +
         geom_text(data = subset(cod.contributions, Year3 == 2013 & Sex2 == "Male" & Cause.of.death == "Cardiovascular"), aes(label = stabbrs), check_overlap = T, size = 2.5)
)

#need to figure out what is weird about this one in NM in 2002 -- something is very odd. 
# ggplotly(ggplot(subset(cod.contributions, Sex2 == "Male" & Cause.of.death == "Cardiovascular"), aes(x = Year3, y = C_x_COD_proportion)) + 
#          geom_line(aes(col = State2)) + facet_wrap(~ Census_Region) +
#          ylab("Contribution to LE Gap (years)") +
#          xlab("Year") +
#          geom_text(data = subset(cod.contributions, Year3 == 2013 & Sex2 == "Male" & Cause.of.death == "Cardiovascular"), aes(label = stabbrs), check_overlap = T, size = 2.5)
# )
```

**Figure 7c: Contribution of communicable disease to the black-white life expectancy gap for males between 1969 and 2013 by US state, in years**

```{r figure7c, fig.align="center", fig.height=8, fig.width=9, echo=F}
ggplotly(ggplot(subset(cod.contributions, Sex2 == "Male" & Cause.of.death == "Communicable"), aes(x = Year3, y = C_x_COD)) + 
         geom_line(aes(col = State2)) + facet_wrap(~ Census_Region) +
         ylab("Contribution to LE Gap (years)") +
         xlab("Year") +
         geom_text(data = subset(cod.contributions, Year3 == 2013 & Sex2 == "Male" & Cause.of.death == "Communicable"), aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

**Figure 7d: Contribution of non-communicable disease to the black-white life expectancy gap for males between 1969 and 2013 by US state, in years**

```{r figure7d, fig.align="center", fig.height=8, fig.width=9, echo=F}
ggplotly(ggplot(subset(cod.contributions, Sex2 == "Male" & Cause.of.death == "Non-communicable"), aes(x = Year3, y = C_x_COD)) + 
         geom_line(aes(col = State2)) + facet_wrap(~ Census_Region) +
         ylab("Contribution to LE Gap (years)") +
         xlab("Year") +
         geom_text(data = subset(cod.contributions, Year3 == 2013 & Sex2 == "Male" & Cause.of.death == "Non-communicable"), aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

**Figure 7e: Contribution of injuries to the black-white life expectancy gap for males between 1969 and 2013 by US state, in years**

```{r figure7e, fig.align="center", fig.height=8, fig.width=9, echo=F}
ggplotly(ggplot(subset(cod.contributions, Sex2 == "Male" & Cause.of.death == "Injuries"), aes(x = Year3, y = C_x_COD)) + 
         geom_line(aes(col = State2)) + facet_wrap(~ Census_Region) +
         ylab("Contribution to LE Gap (years)") +
         xlab("Year") +
         geom_text(data = subset(cod.contributions, Year3 == 2013 & Sex2 == "Male" & Cause.of.death == "Injuries"), aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

**Figure 7f: Contribution of all other causes to the black-white life expectancy gap for males between 1969 and 2013 by US state, in years**

```{r figure7f, fig.align="center", fig.height=8, fig.width=9, echo=F}
ggplotly(ggplot(subset(cod.contributions, Sex2 == "Male" & Cause.of.death == "All other causes"), aes(x = Year3, y = C_x_COD)) + 
         geom_line(aes(col = State2)) + facet_wrap(~ Census_Region) +
         ylab("Contribution to LE Gap (years)") +
         xlab("Year") +
         geom_text(data = subset(cod.contributions, Year3 == 2013 & Sex2 == "Male" & Cause.of.death == "All other causes"), aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

**Figure 7g: Different way to portray the above data for males**

```{r figure7g, fig.align="center", fig.height=8, fig.width=9, echo=F}
for (cod in levels(cod.contributions$Cause.of.death)) {
  print(ggplot(subset(cod.contributions, Sex2 == "Male" & Cause.of.death == cod), aes(x = Year3, y = State2)) + 
          geom_tile(aes(fill = C_x_COD), col = "white") +
          facet_wrap(~ Census_Region, scales = "free_y") + 
          scale_fill_viridis() + theme_minimal() + 
          ylab("Contribution to LE Gap (years)") +
          xlab("Year") + ggtitle(cod))
}
```

**Figure 8a: Contribution of cancer to the black-white life expectancy gap for females between 1969 and 2013 by US state, in years**

```{r figure8a, fig.align="center", fig.height=8, fig.width=9, echo=F}
ggplotly(ggplot(subset(cod.contributions, Sex2 == "Female" & Cause.of.death == "Cancers"), aes(x = Year3, y = C_x_COD)) + 
         geom_line(aes(col = State2)) + facet_wrap(~ Census_Region) +
         ylab("Contribution to LE Gap (years)") +
         xlab("Year") +
         geom_text(data = subset(cod.contributions, Year3 == 2013 & Sex2 == "Female" & Cause.of.death == "Cancers"), 
         aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

**Figure 8b: Contribution of cardiovascular disease to the black-white life expectancy gap for females between 1969 and 2013 by US state, in years**

```{r figure8b, fig.align="center", fig.height=8, fig.width=9, echo=F}
ggplotly(ggplot(subset(cod.contributions, Sex2 == "Female" & Cause.of.death == "Cardiovascular"), aes(x = Year3, y = C_x_COD)) + 
         geom_line(aes(col = State2)) + facet_wrap(~ Census_Region) +
         ylab("Contribution to LE Gap (years)") +
         xlab("Year") +
         geom_text(data = subset(cod.contributions, Year3 == 2013 & Sex2 == "Female" & Cause.of.death == "Cardiovascular"), 
         aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

**Figure 8c: Contribution of communicable disease to the black-white life expectancy gap for females between 1969 and 2013 by US state, in years**

```{r figure8c, fig.align="center", fig.height=8, fig.width=9, echo=F}
ggplotly(ggplot(subset(cod.contributions, Sex2 == "Female" & Cause.of.death == "Communicable"), aes(x = Year3, y = C_x_COD)) + 
         geom_line(aes(col = State2)) + facet_wrap(~ Census_Region) +
         ylab("Contribution to LE Gap (years)") +
         xlab("Year") +
         geom_text(data = subset(cod.contributions, Year3 == 2013 & Sex2 == "Female" & Cause.of.death == "Communicable"), 
         aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

**Figure 8d: Contribution of non-communicable disease to the black-white life expectancy gap for females between 1969 and 2013 by US state, in years**

```{r figure8d, fig.align="center", fig.height=8, fig.width=9, echo=F}
ggplotly(ggplot(subset(cod.contributions, Sex2 == "Female" & Cause.of.death == "Non-communicable"), aes(x = Year3, y = C_x_COD)) + 
         geom_line(aes(col = State2)) + facet_wrap(~ Census_Region) +
         ylab("Contribution to LE Gap (years)") +
         xlab("Year") +
         geom_text(data = subset(cod.contributions, Year3 == 2013 & Sex2 == "Female" & Cause.of.death == "Non-Communicable"), 
         aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

**Figure 8e: Contribution of injuries to the black-white life expectancy gap for females between 1969 and 2013 by US state, in years**

```{r figure8e, fig.align="center", fig.height=8, fig.width=9, echo=F}
ggplotly(ggplot(subset(cod.contributions, Sex2 == "Female" & Cause.of.death == "Injuries"), aes(x = Year3, y = C_x_COD)) + 
         geom_line(aes(col = State2)) + facet_wrap(~ Census_Region) +
         ylab("Contribution to LE Gap (years)") +
         xlab("Year") +
         geom_text(data = subset(cod.contributions, Year3 == 2013 & Sex2 == "Female" & Cause.of.death == "Injuries"), 
         aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

**Figure 8f: Contribution of all other causes to the black-white life expectancy gap for females between 1969 and 2013 by US state, in years**

```{r figure8f, fig.align="center", fig.height=8, fig.width=9, echo=F}
ggplotly(ggplot(subset(cod.contributions, Sex2 == "Female" & Cause.of.death == "All other causes"), aes(x = Year3, y = C_x_COD)) + 
         geom_line(aes(col = State2)) + facet_wrap(~ Census_Region) +
         ylab("Contribution to LE Gap (years)") +
         xlab("Year") +
         geom_text(data = subset(cod.contributions, Year3 == 2013 & Sex2 == "Female" & Cause.of.death == "All other causes"), 
         aes(label = stabbrs), check_overlap = T, size = 2.5)
)
```

**Figure 8g: Different way to portray the above data for females**

```{r figure8g, fig.align="center", fig.height=8, fig.width=9, echo=F}
for (cod in levels(cod.contributions$Cause.of.death)) {
  print(ggplot(subset(cod.contributions, Sex2 == "Female" & Cause.of.death == cod), aes(x = Year3, y = State2)) + 
          geom_tile(aes(fill = C_x_COD), col = "white") +
          facet_wrap(~ Census_Region, scales = "free_y") + 
          scale_fill_viridis() + theme_minimal() + 
          ylab("Contribution to LE Gap (years)") +
          xlab("Year") + ggtitle(cod))
}
```

-add CIs when I have them
